Jug: Software for Reproducible Computation

### 2011-11-23 paper.tex
TXT More on the design choices

\section{Jug}
\subsection{Task Based Architecture}

@@

The status command prints out a summary of the status of all of the tasks.
Figure~\ref{fig:jug-status-output} shows the output of this command using the
example jugfile above. We assume that the jugfile was called \code{images.py}
on disk and that there were 20~images in the directory. We can see that there
are 20~tasks ready to run, while the \code{mean} task is still waiting for the
results of the other tasks.

\subsection{Backends}

@@

The filesystem can support all of the above operations if coded correctly to
avoid race conditions. This is the default backend, identified simply by a
directory name. Inside this directory, files named by a hexadecimal
representation of their hashes. Objects are saved using Python's \code{pickle}
module with zlib compression. As a special case, \code{numpy} arrays are saved
to disk directly. For functionality, there was no need for this inelegant
special case, but \code{numpy} arrays are a very common data type in scientific
programming and saving them directly allows for very fast saving and loading
(they are represented on disk as a header followed by the binary information
they contain).

@@

Results can be obtained from jug in two ways: (1) One can simply write a task
that outputs the desired format. (2) They can be inspected interactively using
the \emph{shell} subcommand.

@@

>>>>

\section{Design Choices}
\subsection{Task Based Architecture}

@@

The status command prints out a summary of the status of all of the tasks.
Figure~\ref{fig:jug-status-output} shows the output of this command using the
example jugfile above. We assume that the jugfile was called \code{images.py}
on disk and that there were 20~images in the directory. We can see that there
are 20~tasks ready to run, while the \code{mean} task is still waiting for the
results of the other tasks.
\subsection{Backends}

@@

The filesystem can support all of the above operations if the backend is coded
correctly to avoid race conditions. This is the default backend, identified
simply by a directory name. Inside this directory, files named by a hexadecimal
representation of their hashes. Objects are saved using Python's \code{pickle}
module with zlib compression. As a special case, \code{numpy} arrays
\citep{numpy} are saved to disk directly. For functionality, there was no need
for this inelegant special case as \code{numpy} can be saved as a pickle dump,
but \code{numpy} arrays are a very common data type in scientific programming
and saving them directly allows for very fast saving and loading (they are
represented on disk as a header followed by the binary information they
contain).

@@

\subsection{Asynchronous Function}

Parallelisation is achieved by running more than one jug process
simmultaneously. All of the synchronisation is outsourced to the backend. This
has certain limitations, but it makes jug usable in cluster computing
environments with a batch architecture. As long as all of the processes can
access the backend, there is no need for them to communicate directly.
Therefore, it can even be the case that processors start working on tasks in
mid-processing.

\subsection{Development Support}

The output of computation can be obtained from jug in two ways: (1) One can
simply write a task that outputs the desired format. (2) They can be inspected
interactively using the \emph{shell} subcommand. It is expected that the first
option will be used for the final version of the computation, whilst the second
one is most helpful during development.

@@

\subsection{Result Invalidation}

It is often the case that one improves a step in a computational pipeline and
wished to obtain the result of recomputing the pipeline while reusing as much
of the previously computed intermediate results. With the manual management of
results I found this to be the most error-prone operation: by not removing
enough of the intermediate files, it was easy to generate an irreproducible
state where part of the computations were made using one version of the code
and part with another.

Therefore, jug adds some support for result invalidation. When a results from a
task are invalidated, all tasks which, directly or indirectly, depend on them
are also marked for deletion.

It is still necessary to manually invalidate tasks. An alternative, which I
considered, was to have the code which implements the task be taken into
account while computing the hash. While this would add another layer of
protection, it would still be possible to make mistakes. If the function
depended on other functions, especially if this was done in a dynamic way, it
could be hard to discover all dependencies (cf.\ sumatra~\citep{sumatra}).
Additionally, minute refactoring of the code would lead to over eager
recomputation. This could make the developer wary of making improvements in
their code, resulting in overall worse code.

Therefore, I chose to force the user to explicitly invalidate his results,
while supporting automatic dependency discovery to make it easier for the user.
